Improving reproducibility of scientific software using Nix/NixOS: A case study on the preCICE ecosystem

### 2024-08-14 paper/paper.tex
style: simplify phase description in section 2

As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
Inputs of these functions are parameters such as the package version, the location of the source code and its dependencies.
Nix then offers built-in functions for common buildsystems and languages such as CMake, Rust, Go and python by implementing different phases.
There are a wide range of phases that always get executed in the same order and package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working build.
The most common phases to override are the \texttt{configurePhase}, which is mostly used for CMake or Autotools configuration, the \texttt{buildPhase}, \texttt{installPhase} and the \texttt{checkPhase} which can for example be used to run unit tests.
It is also worth mentioning that most phases have a \texttt{pre} and \texttt{post} phases, which users can use to make slight adjustments when they need to override existing packages.
Evaluating these functions with the same inputs yields the same outputs.
This is a critical factor for Nix's reproducibility.
Outputs, i.e. the build artifacts, never change after being built once.
Nix builds are, moreover, sandboxed, meaning that there is no internet access possible during a build.

>>>>

As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
Inputs of these functions are parameters such as the package version, the location of the source code and its dependencies.
Nix then offers built-in functions for common buildsystems and languages such as CMake, Rust, Go and python by implementing different phases.
Phases are bash scripts and are already predefined for each buildsystem and always execute in the same order.
Package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working package.
The most important phases are the \texttt{configurePhase}, which is mostly used for buildsystem configuration.
The \texttt{buildPhase} defines the build steps e.g., calling \texttt{make} or \texttt{cargo build}.
Afterwards the \texttt{installPhase} is executed which copies the results into the output directory.
Additionally a \texttt{checkPhase} can be enabled which can be used to run unit tests to insure that the package works as expected.
It is also worth mentioning that most phases have a \texttt{pre} and \texttt{post} phases, which users can use to make slight adjustments when they need to override existing packages.
Evaluating these functions with the same inputs yields the same outputs.
This is a critical factor for Nix's reproducibility.
Outputs, i.e. the build artifacts, never change after being built once.
Nix builds are, moreover, sandboxed, meaning that there is no internet access possible during a build.
